\section{Conclusion}
\subsection{Summary of the Findings}
In this report, we have explored the impact of international fluctuations in gas prices on the Australian
domestic electricity market through the use of a multivariate autoregressive model of the Australian electricity
market, modelled the volatility of global oil and gas markets using a GARCH(1, 1) model in order to quantify
the impacts of the Russian war in Ukraine via a counterfactual analysis, and have examined the possibility of
a cointegrating relationship between global oil and gas spot prices.
\medskip

From these separate but related analyses, we have determined that international fluctuation in gas prices
should have a minimal impact on the National Electricity Market spot price, as there appears to be only a
moderate correlation (at best) between the aggregate gas price within the NEM and the NEM spot price,
(solar and hydro prices appears to be far more correlated with the NEM spot price). The modelling
performed on the international gas and oil markets has quantified the significant impact that the Russian
war in Ukraine has had upon the level of volatility within global oil and gas markets, with oil prices appearing
to become far more volatile as a result of (and, thus, sensitive to) the war in Ukraine. The final analysis on the
possibility of a cointegrating relationship between global oil and gas spot prices disproved the hypothesis
that there exists a cointegrating relationship between these two time series.
\subsection{Implications}
The crucial implication of these findings is that, from both an Australian and international policy perspective,
sanctions on oil appear to have a far greater impact on both domestic and international markets than
sanctions on gas (assuming these sanctions are levelled at a significant producer of both of these
commodities). They highlight the fact that oil appears to be far more sensitive to demand shocks than gas in
the Australian context, with domestic prices of oil (i.e., petrol) becoming more volatile than domestic prices
of gas (i.e., the NEM spot price).
\subsection{Future Research}
There are several future avenues of research that should be examined for a greater understanding of this
topic. Research on the impact of changes to international oil prices on the domestic Australian price of
electricity would allow for a more complete comparison of the possible impacts that sanctions on
hydrocarbons have on the Australian Electricity Market. Additionally, tests for the level of cointegration
between global oil and gas prices on their domestic Australian counterparts (the NEM spot price and
national aggregate petrol prices, respectively) may reveal underlying relationships that were not considered
in this report.